---
title: "Child Health and Development Study"
output: html_document
---

```{r setup, include=FALSE}
# set chunk options so that code is not shown in report
library(knitr)
opts_chunk$set(echo = FALSE)
# load packages here to make dependencies clear

```

## Introduction

This report analyses data from a Child Health and Development Study 
conducted in the US 1960-1967.

```{r import-data}
# add code here to import infant.xlsx

```

The data are a subset of the full study corresponding to live births of a 
single male foetus. Here we will focus on characteristics of the mother 
and child.

```{r select-variables}
# remove redundant variables and variables related to father

```

We exclude extremely premature babies (gestation less than 28 
weeks) and extremely late babies (gestation more than 52 weeks) from the 
analysis to focus on the typical gestation range.

```{r filter-variables}
# filter to exclude extremely premature babies
infant <- filter(infant, gestation > 28 *7 & gestation < 52 * 7)
```

```{r mutate-variables}
# if `ht` is 999, replace it with NA

```

```{r additional-processing}
# processing of other variables ready for later analysis
infant <- infant %>%
    # exclude observations where smoking status unknown
    # also exclude race = 10 as undefined in metadata
    filter(smoke != 9 & !race %in% c(10, 99)) %>% 
    # replace missing codes with NA; create factor from smoking variable
    mutate(gestation = ifelse(gestation == 999, NA, gestation),
           age = ifelse(age == 99, NA, age),
           wt = ifelse(wt == 999, NA, wt),
           race = recode_factor(race, `6` = "Latino", `7` = "Black", 
                                `8` = "Asian", `9` = "Mixed", 
                                .default = "White"),
           smoke = recode_factor(smoke, 
                                 `1` = "currently", `2`= "until pregnancy", 
                                 `3` = "used to", `0` = "never"),
           time = recode_factor(time, `0` = "never smoked",
                                `1` = "still smokes",
                                `2` = "during current preg",
                                `3` = "within 1 yr",
                                `4` = "1 to 2 yrs ago",
                                `5` = "2 to 3 yrs ago",
                                `6` = "3 to 4 yrs ago",
                                `7` = "5 to 9 yrs ago",
                                `8` = "10+ yrs ago",
                                `9` = "quit and don't know"),
           number = recode_factor(number, `0` = "never",
                                  `1` = "1-4",
                                  `2` = "5-9",
                                  `3` = "10-14",
                                  `4` = "15-19",
                                  `5` = "20-29",
                                  `6` = "30-39",
                                  `7` = "40-60",
                                  `8` = "60+",
                                  `9` = "smoke but don't know"))

```


## Low Birth Weight 

An infant is categorised as low weight if its birth weight is ≤ 2500 
grams, regardless of gestation. The following table shows the distribution 
of birth weight in the sample

```{r table-bwt}
# write an analysis pipe to summarise the number of infants in 
# different weight categories
tab <- 
    
kable(tab)
```

### ggplot2 

The following plot shows the association between mother's age and birth weight depending on smoking status. 

```{r ggplot-bwt-v-age}
# use ggplot to inspect the association between mother's age and birth

```

### linear model

Based on exploratory analysis, we fit a baseline model for birth weight 
that includes gestation and key maternal characteristics, excluding age

```{r}
# use lm to  model birth weight against both gestation and the key
# maternal characteristics weight (`wt`), height (`ht`), parity and race

```

We then add the smoking variables to determine if they have a 
significant effect on age over and above the baseline variables

```{r}
# use lm to  model birth weight against both gestation and the key
# maternal characteristics weight (`wt`), height (`ht`), parity and race

```

The following ANOVA table compares the two models

```{r}
# use anove to compare models

```


